\chapter{Discussion}

\section{About the project}\label{about-the-project}

This internship has led to a functional device that can be used to explore graphs; it has a flexible software application that has been built considering further collaboration and accepting other interaction techniques.

By an iterative process, we have managed to outline some main criteria for usable tactons. We did manage to finish a user study giving some first clues on (<<<<>>>>)

Since the analysis on the first user study is still running, it is risky to say that the interaction G is overall better than the M. Still, every subject has preferred the latter in spite of a lack of significant improvement in effectiveness. The idea of having a guidance may be very relieving.

Many positive remarks have emerged from this first user study. A longer training to become at complete ease was a popular remark (<<<<reformulate, not clear>>>). They have also shared the fact that they found it more difficult to follow diagonals than horizontal or vertical lines. 

Even though the project has been delayed by some optimisations, I think the results are appropriate (<<< sufficient?>>>) given the time frame.


\section{Acquired skills}\label{acquired-skills}

This internship has implicated me in various domain of expertise. I have applied a user-centered development while still gathering experience in technical skills which I am grateful for.

I had the chance to work in two different research groups from two different countries which is definitely a unique opportunity for a non-PhD student.

Being supervised by a neuroscientist taught me another approach in science.

Finally, 

This a non exhaustive list of the technology I have used ordered by the time I have spent on it:

\begin{itemize}
	\item Python
	\item Arduino
	\item TUIO protocol
	\item Tkinter framework
	\item Git
	\item R (for statistics)
	\item Socket
	\item reStructuredText for Python documentation
\end{itemize}

From the perspective of HCI skills I have learned:

\begin{itemize}
	\item to build rapid prototypes for tactile signals
	\item to conduct a background research
	\item to include the users' perspective even when resources are scarce
	\item to build, try and conduct a user study
	\item to analyse data resulting in such study
	\item to iterate over prototypes by design walkthrough and inject new ideas
	\item to build evolutive software
	\item to adapt a user-centered process for a specific deficiency
\end{itemize}


